### 2023-03-02 paper/jmlda/jmlda-template-eng.tex

\documentclass[12pt, twoside]{article}
\usepackage{jmlda}
\usepackage[utf8]{inputenc}
\usepackage[english,main=russian]{babel}
\usepackage[T2A]{fontenc}
\newcommand{\hdir}{.}

\begin{document}
\English

\title
	[JMLDA paper template] % short title for page headings, not necessary if a full title fits the headings
    {Machine Learning and Data Analysis journal paper template} % full title
\author
	[F.\,S.~Author] % short list of the authors (<= 3) for page headings, is necessary only if the full list does not fit the headings
	{F.\,S.~Author, F.\,S.~Co-Author, and F.\,S.~Name} % full list of the authors, presented in the table of contetns of the issue
    [F.\,S.~Author$^1$, F.\,S.~Co-Author$^2$, and F.\,S.~Name$^{1, 2}$] % list of the authors presented in the title page of the article, is necessary only if it differs from the full list of the authors in braces, i.e. '{' and '}'
\email
    {author@site.ru; co-author@site.ru; co-author@site.ru}
\thanks
    {The research was
     %partially
     supported by the Russian Foundation for Basic Research (grants 00-00-0000 and 00-00-00001).}
\organization
    {$^1$Organization, address;
     $^2$Organization, address}
\abstract
    {This is the template of the paper submitted to the journal ``Machine Learning and Data Analysis''.

	\noindent
	The title should be concise and informative. Titles are often used in information-retrieval systems. Avoid abbreviations and formulae where possible.
	Please clearly indicate the last names and initials of each author and check that all names are accurately spelled. Present the authors' affiliation addresses where the actual work was done.
	Provide the full postal address of each affiliation, including the country name and, if available, the
	e-mail address of each author.
	Provide only institutional affiliation, department/division affiliation are not required.

	\noindent
	A concise and factual abstract is required.
	The purpose of the abstract is to provide a summary~of the paper enabling the reader to decide whether or not to read the full text.
    	The abstract should state briefly the purpose of the research, the principal results and major conclusions.
    	An abstract is often presented separately from the article, so it must be able to stand alone.
    	For this reason, References should be avoided, but if essential, then cite the author(s) and year(s).
    	Also, non-standard or uncommon abbreviations should be avoided, but if essential they must be defined at their first mention in the abstract itself.
    	The requirements on the size of the abstract is about 200--300 words.
    	It should be provided in the next structured manner:

	\noindent
	\textbf{Background}:	One paragraph about the problem, existent approaches and its limitations.

	\noindent
	\textbf{Methods}: One paragraph about proposed method and its novelty.

	\noindent
	\textbf{Results}: One paragraph about major properties of the proposed method and experiment results if applicable.

	\noindent
	\textbf{Concluding Remarks}: One paragraph about the place of the proposed method among existent approaches.

	\noindent
	Immediately after the abstract, provide 5-7 keywords, avoiding general and plural terms and multiple concepts (avoid, for example, ``and'', ``of'').
	Use keywords that are specific and that reflect what is essential about the paper.
	Use keywords from the abstract, introduction and conclusion.
	These keywords will be used for indexing purposes.

	\noindent
    	\textbf{Keywords}: \emph{keyword; keyword; more keywords, separated by ``;''}}

\titleRus
    [Шаблон статьи для публикации] % short title for page headings, not necessary if a full title fits the headings
    {Шаблон статьи для публикации в журнале <<Машинное обучение и анализ данных>>} % full title
\authorRus
    [И.\,О.~Автор] % short list of the authors (<= 3) for page headings, is necessary only if the full list does not fit the headings
    {И.\,О.~Автор, И.\,О.~Соавтор, И.\,О.~Фамилия} % full list of the authors, presented in the table of contetns of the issue
    [И.\,О.~Автор$^1$, И.\,О.~Соавтор$^2$, И.\,О.~Фамилия$^{1,2}$] % list of the authors presented in the title page of the article, is necessary only if it differs from the full list of the authors in braces, i.e. '{' and '}'
\thanksRus
    {Работа выполнена при
     %частичной
     финансовой поддержке РФФИ, проекты \No\ \No 00-00-00000 и 00-00-00001.}
\organizationRus
    {$^1$Организация, адрес; $^2$Организация, адрес}
\abstractRus
    {Данный текст является шаблоном статьи, подаваемой для публикации в журнале <<Машинное обучение и анализ данных>>.

    Аннотация описывает основную цель работы,
    особенности предлагаемого подхода и~основные результаты.
    Сведения, содержащиеся в заглавии статьи, не должны повторяться в тексте авторского резюме.
    В аннотации не должно быть ссылок на литературу и, по возможности, формул.

	Также необходимо представить расширенную структурированную аннотацию на английском языке объемом 200--300 слов.	
	Английская аннотация может не быть дословным переводом русского текста и должна быть написана хорошим английским языком.

	В титульном заголовке необходимо указать полный, официально принятый, переводной вариант названия организации.
	Указывать нужно только ту часть названия, которая относится к понятию юридического лица,
	не вписывая названий кафедры, лаборатории или другого структурного подразделения внутри организации.
	Необходимо указать полный юридический адрес, или, как минимум, город и страну.

 	При выборе ключевых слов основным критерием является их потенциальная ценность для выражения содержания документа или для его поиска.
	В качестве ключевых слов могут использоваться термины из названия, аннотации, вступительной и заключительной части текста статьи.
 	При подборе ключевых слов рекомендуется использовать базовые термины вместе с более сложными, допускается использование повторов и синонимов.
	Не рекомендуется использование слишком сложных слов, слов в кавычках, слов с запятыми.
	По возможности следует применять слова в основной форме именительного падежа единственного числа.
	Рекомендуемое количество ключевых слов~-- 5-7, количество слов внутри ключевой фразы~-- не более 3.

\bigskip
\noindent
\textbf{Ключевые слова}: \emph {ключевое слово; ключевое слово; еще ключевые слова, разделенные <<;>>}
}

%these fields are filled in by the journal editors
\doi{10.21469/22233792}
\receivedRus{01.01.2017}
\receivedEng{January 01, 2017}

\maketitle
\linenumbers

\section{Introduction}
\noindent %this command is placed at the beginning of the first sentence of each paragraph/section only.
In this section state the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results.

\section{Preparing a manuscript}
\noindent
Manuscripts are prepared using \verb'jmlda.sty' style package.
You are recommended to use \verb'jmlda_rus.bst' and \verb'jmlda_eng.bst' style files for generating bibliography using Bib\TeX.

Visit the \url{http://jmlda.org/?lang=en} website for detailed submission instructions, templates and other information.

Please note that this file must be saved in~\verb'UTF-8' encoding. Where possible select~\verb'UTF-8 without BOM' encoding. 
To change the encoding please use \verb'Sublime Text' or \verb'Notepad++' text editors.

\section{Structure of the article}
\noindent
Divide your article into clearly defined and numbered sections and paragraphs.

\paragraph{Paragraph}
\noindent
Sections and paragraphs are numbered and have a brief heading.

%please do not change the name of this section if it is present
\section{Concluding Remarks}
This section should provide the summary and explore the significance of the results achieved and list problems not yet solved.
Results should be clear and concise.

%%%% please specify doi of the cited item if possible, see~\bibitem{article}
%%%% Crossref doi of the item can be retrieved at http://www.crossref.org/guestquery/
\begin{thebibliography}{99}

\bibitem{book}
	\BibAuthor{Goossens,~M., F. Mittelbach, and A.~Samarin}. 1994.
	\BibTitle{The \LaTeX\ companion}.
	2nd ed.
	Reading, MA: Addison-Wesley. 528 p.

\bibitem{article}
	\BibAuthor{Zagurenko,~A.\,G., V.\,A.~Korotovskikh, A.\,A.~Kolesnikov, A.\,V.~Timonov, and D.\,V.~Kardymon}. 2008.
	Tekhniko-ekonomicheskaya optimizatsiya dizayna gidrorazryva plasta
	[Technical and economic optimization of the design of hydraulic fracturing].
	\BibJournal{Neftyanoe Khozyaystvo} [Oil Industry] 11(1):54--57.
	\BibDoi{10.3114/S187007708007}. (In Russian)

\bibitem{webArticle}
	\BibAuthor{Blaga,~P.\,A.} 2007.
	Commutative Diagrams with XY-pic II. Frames and Matrices.
	\BibJournal{PracTEX J.}  4.
	Available at: \BibUrl{https://tug.org/pracjourn/2007-1/blaga/blaga.pdf}
    (accessed February 20, 2007).

\bibitem{webResource}
	XYpic.
	Available at: \BibUrl{http://akagi.ms.u-tokyo.ac.jp/input9.pdf}
	(accessed April 09, 2015).

\bibitem{inproceedingsRus}
	\BibAuthor{Usmanov,~T.\,S., A.\,A.~Gusmanov, I.\,Z.~Mullagalin, R.\,Yu.~Mukhametshina, A.\,N.~Chervyakova, and A.\,V.~Sveshnikov.} 2007.
	Osobennosti proektirovaniya razrabotki mestorozhdeniy s primeneniem gidrorazryva plasta
	[Features of the design of field development with the use of hydraulic fracturing].
	\BibJournal{6th Symposium (International) ``New Energy Saving Subsoil Technologies and the
	Increasing of the Oil and Gas Impact'' Proceedings}.
	Moscow:~Publisher. 267--272. (In Russian)

\bibitem{inproceedingsEng}
    \BibAuthor{Author,~N.} 2009.
    Paper title.
    \BibJournal{10th Conference (International) on Any Science Proceedings}.
    Place of publication: Publisher. 111--122.

\bibitem{techreport}
	\BibAuthor{Lambert,~P.} 1993.
  	\BibTitle{The title of the work}.
  	Place of publication:~The institution that published.  Report~2.

\end{thebibliography}

\maketitleSecondary
\Russian
%%%% please specify doi of the cited item if possible, see~\bibitem{article}
%%%% Crossref doi of the item can be retrieved at http://www.crossref.org/guestquery/
\begin{thebibliography}{99}
\bibitem{book}
    \BibAuthor{Гуссенс~М., Миттельбах~Ф., Cамарин~А.}
    \BibTitle{Путеводитель по пакету \LaTeX\ и~его расширению \LaTeXe} / Пер. с англ.~---
    М.:~Мир, 1999. 606~с.
    (\BibAuthor{Goossens M., Mittelbach F., Samarin A.}
     \BibTitle{The \LaTeX\ companion}.~--- 2nd ed.~--- Reading, MA, USA: Addison-Wesley, 1994. 528 p.)

\bibitem{article}
    \BibAuthor{Загуренко~А.\,Г., Коротовских~В.\,А., Колесников~А.\,А., Тимонов~А.\,В., Кардымов~Д.\,В.}
    Технико-экономическая оптимизация дизайна гидроразрыва пласта~//
    \BibJournal{Нефтяное хозяйство}, 2008. Т.~11. \No\,1. С.~54--57.
	\BibDoi{10.3114/S187007708007}.

\bibitem{webArticle}
	\BibAuthor{Blaga~P.\,A.}
	Commutative Diagrams with XY-pic II. Frames and Matrices~//
	\BibJournal{PracTEX J.}, 2007. Vol.\,4.
	URL: \BibUrl{https://tug.org/pracjourn/2007-1/blaga/blaga.pdf}.

\bibitem{webResource}
	XYpic.
	URL: \BibUrl{http://akagi.ms.u-tokyo.ac.jp/input9.pdf}.

\bibitem{inproceedingsRus}
	\BibAuthor{Усманов~Т.\,С., Гусманов~А.\,А., Муллагалин~И.\,З., Мухаметшина~Р.\,Ю., Червякова~А.\,Н., Свешников~А.\,В.}
	Особенности проектирования разработки месторождений с применением гидроразрыва пласта~//
	\BibJournal{Труды 6-го Междунар. симп. <<Новые ресурсосберегающие технологии недропользования и повышения нефтегазоотдачи>>}.~---
	М.:~Издательство, 2007. С.~267--272.

\bibitem{inproceedingsEng}
    \BibAuthor{Author~N.}
    Paper title~//
    \BibJournal{10th Conference (International) on Any Science Proceedings}.~---
    Place of publication: Publisher, 2009. P.~111--122.

\bibitem{techreport}
	\BibAuthor{Lambert~P.}
  	\BibTitle{The title of the work}.
  	Place of publication:~The institution that published, 1993.  Report~2.

\end{thebibliography}

\end{document}